---
title: "Statistical Model"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

```{r message=FALSE, echo=FALSE, warning=FALSE}
library(dplyr)
library(readxl)
library(knitr)
library(kableExtra)
library(ggplot2)
library(gridExtra)
library(patchwork)
library(pROC)
library(rpart)
library(rpart.plot)
library(randomForest)
library(gbm)
library(caret)
library(xgboost)
library(plotly)
library(leaflet)
library(RColorBrewer)

# URLs for the datasets
urls <- c(
  "https://raw.githubusercontent.com/KaiwenYangUT/Robbery_Crime_Classification/main/data/Auto_Theft_Open_Data.csv",
  "https://raw.githubusercontent.com/KaiwenYangUT/Robbery_Crime_Classification/main/data/Bicycle_Thefts_Open_Data.csv",
  "https://raw.githubusercontent.com/KaiwenYangUT/Robbery_Crime_Classification/main/data/Break_and_Enter_Open_Data.csv",
  "https://raw.githubusercontent.com/KaiwenYangUT/Robbery_Crime_Classification/main/data/Homicides_Open_Data_(ASR-RC-TBL-002).csv",
  "https://raw.githubusercontent.com/KaiwenYangUT/Robbery_Crime_Classification/main/data/Robbery_Open_Data.csv",
  "https://raw.githubusercontent.com/KaiwenYangUT/Robbery_Crime_Classification/main/data/Shooting_and_Firearm_Discharges_Open_Data.csv",
  "https://raw.githubusercontent.com/KaiwenYangUT/Robbery_Crime_Classification/main/data/Theft_Over_Open_Data.csv"
)

# Function to read and preprocess data
read_and_process <- function(url) {
  data <- read.csv(url)
  return(data)
} 

# Read and preprocess all datasets
datasets <- lapply(urls, read_and_process)

# Combine datasets into one
crime <- bind_rows(datasets, .id = "Dataset")

# URL of the Excel file on GitHub
github_url <- "https://github.com/KaiwenYangUT/Robbery_Crime_Classification/raw/main/data/Income_Dataset.xlsx"

# File destination on your local machine
local_file <- "Income_Dataset.xlsx"

# Download the file from GitHub
download.file(github_url, local_file, mode = "wb")

# Read the Excel file
income <- read_excel(local_file)

# - Create a indicator variable Robbery
crime <- crime %>%
  mutate(Robbery = MCI_CATEGORY == "Robbery")


# - Transform OCC_MONTH into int
# Create new column OCC_Month_Numeric that transform chr OCC_MONTH into int.
crime$OCC_Month_Numeric <- match(crime$OCC_MONTH, month.name)


# - Create a new column Occurence_Time with the transformation of OCC_YEAR, OCC_MONTH, OCC_DAY and OCC_HOUR.
crime$Occurrence_Time <- as.POSIXct(paste(crime$OCC_YEAR, crime$OCC_Month_Numeric, crime$OCC_DAY, crime$OCC_HOUR, sep="-"), format="%Y-%m-%d-%H")


# - Convert Day of the Week to numeric representation
crime$Day_Of_Week_Numeric <- match(weekdays(crime$Occurrence_Time), c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))


# - Keep only necessary columns and rename them
crime <- crime[, c("MCI_CATEGORY", "Robbery", "Occurrence_Time","OCC_YEAR", "OCC_Month_Numeric", "Day_Of_Week_Numeric", "OCC_HOUR", "NEIGHBOURHOOD_158", "LOCATION_TYPE", "PREMISES_TYPE", "LONG_WGS84", "LAT_WGS84")]
colnames(crime) <- c("Crime_Category","Robbery", "Occurrence_Time", "Year", "Month", "Day_Of_Week_Numeric", "Hour", "Neighbourhood", "Location_Type", "Premises_Type", "Longitude", "Latitude")


# - Create a new variable Season by grouping the month
# Create Variable Season and Summary
get_season <- function(month) {
  if (month %in% c(3, 4, 5)) {
    return("Spring")
  } else if (month %in% c(6, 7, 8)) {
    return("Summer")
  } else if (month %in% c(9, 10, 11)) {
    return("Fall")
  } else {
    return("Winter")
  }
}


# Add a new column for season based on Month
crime$Season <- sapply(crime$Month, get_season)


# Define average sunset and sunrise times in Toronto (example times)
average_sunset_time <- 18
average_sunrise_time <- 6


# Create the indicator variable
crime$Darkness <- ifelse(crime$Hour >= average_sunset_time | crime$Hour < average_sunrise_time, 1, 0)


# - Create new dataset crime_2020 by filtering crime happened in the year of 2020
crime_2020 <- subset(crime, Year == 2020)


# - Cleaning the dataset first by removing crime with non-recorded neighbourhood
crime_2020 <- subset(crime_2020, Neighbourhood != "NSA")

income_2020 <- income[, c("Neighbourhood Name",
                          
                        "Median total income in 2020  among recipients ($)",
                        "Average total income in 2020 among recipients ($)", 
                        
                        "Number of persons in private households", 
                        
                        "One-census-family households without additional persons",
                        "Couple-family households",
                        "One-parent-family households",
                        "Multigenerational households",
                        "Multiple-census-family households",
                        "One-census-family households with additional persons",
                        "Two-or-more-person non-census-family households",
                        "Median total income of two-or-more-person households in 2020 ($)",
                        "Average total income of two-or-more-person households in 2020 ($)",

                        "One-person households",
                        "Median total income of one-person households in 2020 ($)",
                        "Average total income of one-person households in 2020 ($)",
                        
                        "Average household size",
                        "Median total income of household in 2020 ($)",
                        "Average total income of household in 2020 ($)",
                        
                        "Total - Private households by tenure - 25% sample data",
                        "Owner",
                        "Renter",
                        
                        "Total - Labour force aged 15 years and over by class of worker including job permanency - 25% sample data",
                        "Permanent position",
                        "48-49 Transportation and warehousing",
                        "62 Health care and social assistance")]

colnames(income_2020) <- c("Neighbourhood", 
                           
                           "Individual_Median_Income", 
                           "Individual_Average_Income",
                           
                           "Population",
                           
                           "One-census-family households without additional persons",
                           "Couple-family households",
                           "One-parent-family households",
                           "Multigenerational households",
                           "Multiple-census-family households",
                           "One-census-family households with additional persons",
                           "Two-or-more-person non-census-family households",
                           "One+_people_Household_Median_Income",
                           "One+_people_Household_Average_Income",
                           
                           "Number_of_One_person_Households",
                           "One_Person_Household_Median_Income",
                           "One_Person_Household_Average_Income",

                           
                           "Average_Household_Size",
                           "Household_Median_Income",
                           "Household_Average_Income",
                           
                           "Number_of_Private_Household_by_Tenure",
                           "Number_of_Owner",
                           "Number_of_Rental_Properties",
                           
                           "Total_Labour_Force",
                           "Permanent_Job_Position",
                           "Transportation_Services_Position", 
                           "Healthcare_Services_Position")

# Create the new column
income_2020$Number_of_One_plus_people_Household_Median_Income <- rowSums(income_2020[, c(
  "Multiple-census-family households", 
  "One-census-family households without additional persons",
  "Couple-family households",
  "One-parent-family households",
  "Multigenerational households",
  "One-census-family households with additional persons",
  "Two-or-more-person non-census-family households"
)], na.rm = TRUE)

# ratio of permanent job positions / total labour force
income_2020$Permanent_Job_and_Labour_Force_Ratio <- 
  income_2020$Permanent_Job_Position / income_2020$Total_Labour_Force

# ratio of number of transportation service workers / number of population
income_2020$Transportation_Service_Worker_and_Population_Ratio <-
  income_2020$Transportation_Services_Position / income_2020$Population

# ratio of number of health care service workers / number of population
income_2020$HealthCare_Service_Worker_and_Population_Ratio <- 
  income_2020$Healthcare_Services_Position / income_2020$Population

# Difference in individuals' median and average income
income_2020$Difference_in_Individual_Median_Average_Income <- abs(income_2020$Individual_Average_Income - income_2020$Individual_Median_Income)

# Remove the original columns
income_2020 <- income_2020[, !names(income_2020) %in% c(
  "Multiple-census-family households",
  "One-census-family households without additional persons",
  "Couple-family households",
  "One-parent-family households",
  "Multigenerational households",
  "One-census-family households with additional persons",
  "Two-or-more-person non-census-family households"
)]

# Extract the neighborhood columns from each dataset
neighborhood_crime <- crime_2020$Neighbourhood
neighborhood_income <- income_2020$Neighbourhood

# Find the differences
difference_crime_to_income <- setdiff(neighborhood_crime, neighborhood_income)
difference_income_to_crime <- setdiff(neighborhood_income, neighborhood_crime)

# Output the differences
# print("Neighborhoods in crime dataset not present in income dataset:")
# print(difference_crime_to_income)

# print("Neighborhoods in income dataset not present in crime dataset:")
# print(difference_income_to_crime)

# Replace the values
income_2020$Neighbourhood <- gsub("O`Connor Parkview", "O'Connor-Parkview", income_2020$Neighbourhood)
income_2020$Neighbourhood <- gsub("Danforth-East York", "Danforth East York", income_2020$Neighbourhood)
income_2020$Neighbourhood <- gsub("Taylor Massey", "Taylor-Massey", income_2020$Neighbourhood)
income_2020$Neighbourhood <- gsub("East End Danforth", "East End-Danforth", income_2020$Neighbourhood)
income_2020$Neighbourhood <- gsub("Cabbagetown-South St. James Town", "Cabbagetown-South St.James Town", income_2020$Neighbourhood)
income_2020$Neighbourhood <- gsub("North St. James Town", "North St.James Town", income_2020$Neighbourhood)
income_2020$Neighbourhood <- gsub("Yonge-St. Clair", "Yonge-St.Clair", income_2020$Neighbourhood)
income2020 <- na.omit(income_2020)

income2020 <- na.omit(income_2020)
crime_2020 <- na.omit(crime_2020)
merged_data <- merge(crime_2020, income_2020, by = "Neighbourhood", all.x = TRUE)
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
# Assigning values to variables
glm_train_accuracy <- 0.8382214
glm_test_accuracy <- 0.8324639
glm_precision <- 0.6756757
glm_recall <- 0.0294811
glm_f1_score <- 0.0564972

train_accuracy_tree <- 0.8417477
test_accuracy_tree <- 0.8342697
precision_tree <- 0.5916667
recall_tree <- 0.0837264
f1_score_tree <- 0.1466942

train_accuracy_bag <- 0.9980218
test_accuracy_bag <- 0.8733949
precision_bag <- 0.6817420
recall_bag <- 0.4799528
f1_score_bag <- 0.5633218

train_accuracy_rf <- 0.9973338
test_accuracy_rf <- 0.8836276
precision_rf <- 0.7723577
recall_rf <- 0.4481132
f1_score_rf <- 0.5671642

boost_train_accuracy <- 0.8395975
boost_test_accuracy <- 0.8338684
boost_precision <- 0.7631579
boost_recall <- 0.0341981
boost_f1_score <- 0.0654628

xgb_train_accuracy <- 0.8879333
xgb_test_accuracy <- 0.8430979
xgb_precision <- 0.5993976
xgb_recall <- 0.2346698
xgb_f1_score <- 0.3372881

```


```{r message=FALSE, echo=FALSE, warning=FALSE}
merged_data <- merged_data[, c("Neighbourhood", "Robbery", "Month", "Day_Of_Week_Numeric", "Hour",
                               "Premises_Type", "Longitude", "Latitude", "Season", "Darkness",
                               "Individual_Median_Income", "Individual_Average_Income", "Population",
                               "Average_Household_Size", "Number_of_Rental_Properties",
                               "Permanent_Job_and_Labour_Force_Ratio", 
                               "Transportation_Service_Worker_and_Population_Ratio",
                               "HealthCare_Service_Worker_and_Population_Ratio",
                               "Difference_in_Individual_Median_Average_Income"
)]
  
final_data <- subset(merged_data, select = -Neighbourhood)

```
```{r message=FALSE, echo=FALSE, warning=FALSE}
# Split data into training and testing sets
set.seed(123)
train_index <- sample(nrow(final_data), 0.7 * nrow(final_data))
train_data <- final_data[train_index, ]
test_data <- final_data[-train_index, ]

train_data$Robbery <-as.factor(train_data$Robbery)
test_data$Robbery <-as.factor(test_data$Robbery)
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
# - Random Forest

train_data_rf <- train_data
test_data_rf <- test_data 

train_data_rf$Robbery <-as.factor(train_data_rf$Robbery)
test_data_rf$Robbery <-as.factor(test_data_rf$Robbery)
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
# Fit the Random Forest model
robbery_rf <- randomForest(
  Robbery ~ .,
  data = train_data_rf,
  na.action = na.omit
)
```

# Summary of Model Performance

Each model underwent rigorous evaluation, including training accuracy, testing accuracy, precision, recall, and F1 score assessments, providing a comprehensive understanding of model performance. The performance scores are shown in the Performance Metrics of Different Models.

In the table, Bagging has the highest training accuracy, and Random Forest has the second highest training accuracy, which is reasonable due to their design nature. Comparing test accuracy, we can see that both Random Forest and Bagging perform better than other models, achieving accuracies of 0.8836276 and 0.8733949, respectively.

Upon examination of precision, recall, and F1 Score, it is evident that Random Forest and Bagging still exhibit better performance than other models. Random Forest demonstrates comparatively better performance in precision and F1 Score, while Bagging excels in recall. Although Bagging achieves a higher recall score than Random Forest, and recall is an important measure for identifying actual robbery cases, which aligns with our primary objective, Random Forest achieves much higher precision than Bagging, despite the difference in recall for both models. Since our goal is to help people avoid becoming victims of robbery, it is worth noting that higher precision may be more effective in reducing unnecessary panic, as precision measures the proportion of actual robberies among predicted robberies. Consequently, I selected Random Forest as the final model.

```{r message=FALSE, echo=FALSE, warning=FALSE}
# Create a summary table
summary_table <- data.frame(
  Model = c("GLM", "Classification Tree", "Bagging", "Random Forest", "Boosting", "XGBoosting"),
  Train_Accuracy = c(glm_train_accuracy, train_accuracy_tree, train_accuracy_bag, train_accuracy_rf, boost_train_accuracy, xgb_train_accuracy),
  Test_Accuracy = c(glm_test_accuracy, test_accuracy_tree, test_accuracy_bag, test_accuracy_rf, boost_test_accuracy, xgb_test_accuracy),
  Precision = c(glm_precision, precision_tree, precision_bag, precision_rf,boost_precision, xgb_precision),
  Recall = c(glm_recall, recall_tree, recall_bag, recall_rf,boost_recall, xgb_recall),
  F1_Score = c(glm_f1_score, f1_score_tree, f1_score_bag, f1_score_rf, boost_f1_score, xgb_f1_score)
)

# Print the summary table using kable
kable(summary_table, align = c("l", rep("c", ncol(summary_table) - 1)), 
      caption = "Performance Metrics of Different Models")

```

# Final Model and Variable Importance

The variable importance of the fitted Random Forest model is shown in Figure 7 below. Spatial indicators such as Longitude, Latitude, and Premises_Type are ranked as the three most important variables in the model. Following spatial indicators, the second most important factors are temporal factors, including Hour, Month, Day_of_Week, and Season. Finally, we come to the socio-economic factors of neighborhoods, ranked in descending order of importance: Number_of_Rental_Properties, Permanent_Job_and_Labour_Force_Ratio, Transportation_Service_Worker_and_Population_Ratio, HealthCare_Service_Worker_and_Population_Ratio, and so on. It is worth noting that the newly created variable Difference_in_Individual_Median_Average_Income is ranked higher than both Individual_Median_Income and Individual_Average_Income, implying success in variable creation.

```{r message=FALSE, echo=FALSE, warning=FALSE}
vi <- importance(robbery_rf)[, "MeanDecreaseGini"]
vi_sorted <- sort(vi, decreasing = TRUE)
vi_df <- data.frame(Variable = names(vi_sorted), Importance = vi_sorted)

# Create a named vector for renaming
readable_names <- setNames(
  c("Robbery", "Month", "Day of Week", "Hour", "Premises Type", "Longitude", 
    "Latitude", "Season", "Darkness", "Median Individual Income", 
    "Average Individual Income", "Population", "Average Household Size", 
    "Number of Rental Properties", "Job and Labour Force Ratio", 
    "Transportation Worker to Population Ratio", 
    "Healthcare Worker to Population Ratio", 
    "Difference in Median and Average Income"),
  c("Robbery", "Month", "Day_Of_Week_Numeric", "Hour", "Premises_Type", 
    "Longitude", "Latitude", "Season", "Darkness", 
    "Individual_Median_Income", "Individual_Average_Income", "Population", 
    "Average_Household_Size", "Number_of_Rental_Properties", 
    "Permanent_Job_and_Labour_Force_Ratio", 
    "Transportation_Service_Worker_and_Population_Ratio", 
    "HealthCare_Service_Worker_and_Population_Ratio", 
    "Difference_in_Individual_Median_Average_Income")
)

# Replace names in the data frame
vi_df$Variable <- readable_names[vi_df$Variable]



ggplot(vi_df, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Variable Importance", x = "Variables", y = "Importance") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

```