---
title: "Key Variable Statistics"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressPackageStartupMessages({
  library(dplyr)
  library(DT)
  library(plotly)
  library(tidyr)
})
source("scripts/prepare_project_data.R")
project_data <- load_project_data(year = 2020)
merged_data <- project_data$merged
```

<div class="hero-panel">
<div class="hero-grid">
<div class="hero-copy">
<div class="hero-kicker">Variable Overview</div>

# Key Variable Statistics

<p class="hero-lead">This page summarizes the main numerical, temporal, and categorical variables used in the project. The focus is on clear comparison: a compact summary table for continuous variables and interactive distribution charts for seasonal, monthly, weekly, and hourly patterns.</p>
</div>

<div class="hero-meta">
<div class="stat-card">
<span class="stat-value">9</span>
<p class="stat-label">continuous variables summarized in the table below</p>
</div>
<div class="stat-card">
<span class="stat-value">24</span>
<p class="stat-label">hourly bins used to inspect time-of-day concentration</p>
</div>
<div class="stat-card">
<span class="stat-value">7</span>
<p class="stat-label">premise and time variables carried into the visual and modeling workflow</p>
</div>
</div>
</div>
</div>

# Continuous Numerical Variables

```{r continuous-summary}
summary_vars <- merged_data[, c(
  "Longitude",
  "Latitude",
  "Population",
  "Average_Household_Size",
  "Number_of_Rental_Properties",
  "Permanent_Job_and_Labour_Force_Ratio",
  "Transportation_Service_Worker_and_Population_Ratio",
  "HealthCare_Service_Worker_and_Population_Ratio",
  "Difference_in_Individual_Median_Average_Income"
)]

summary_table <- tibble(
  Variable = c(
    "Longitude",
    "Latitude",
    "Population",
    "Average Household Size",
    "Number of Rental Properties",
    "Permanent Job and Labour Force Ratio",
    "Transportation Service Worker and Population Ratio",
    "HealthCare Service Worker and Population Ratio",
    "Difference in Individual Median Average Income"
  ),
  Min = sapply(summary_vars, min, na.rm = TRUE),
  `1st Qu.` = sapply(summary_vars, quantile, probs = 0.25, na.rm = TRUE),
  Median = sapply(summary_vars, median, na.rm = TRUE),
  Mean = sapply(summary_vars, mean, na.rm = TRUE),
  `3rd Qu.` = sapply(summary_vars, quantile, probs = 0.75, na.rm = TRUE),
  Max = sapply(summary_vars, max, na.rm = TRUE)
) %>%
  mutate(
    across(-Variable, ~ ifelse(abs(.) >= 1000, format(round(., 0), big.mark = ","), sprintf("%.3f", .)))
  )

datatable(
  summary_table,
  rownames = FALSE,
  class = "compact stripe hover",
  options = list(
    pageLength = 9,
    dom = "tip",
    autoWidth = TRUE,
    scrollX = TRUE,
    columnDefs = list(
      list(width = "28%", targets = 0),
      list(className = "dt-center", targets = 1:6)
    )
  )
) %>%
  formatStyle(
    columns = names(summary_table),
    backgroundColor = "#fffaf2",
    color = "#2b2620",
    borderColor = "rgba(31, 29, 26, 0.08)"
  ) %>%
  formatStyle(
    0,
    fontWeight = "600"
  )
```

<div class="note-callout">
The largest spread appears in neighborhood population, rental property counts, and the difference between median and average income. Geographic coordinates vary within a narrower band, as expected for a single city.
</div>

# Temporal and Categorical Distributions

```{r discrete-plots}
season_counts <- merged_data %>%
  count(Season) %>%
  mutate(Season = factor(Season, levels = c("Spring", "Summer", "Fall", "Winter")))

month_counts <- merged_data %>%
  count(Month) %>%
  mutate(Month = factor(Month, levels = 1:12, labels = month.abb))

day_counts <- merged_data %>%
  count(Day_Of_Week_Numeric) %>%
  mutate(Day = factor(c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")[Day_Of_Week_Numeric],
                      levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")))

hour_counts <- merged_data %>%
  count(Hour) %>%
  mutate(Hour_Label = sprintf("%02d:00", Hour))

p_season <- plot_ly(
  season_counts,
  x = ~Season,
  y = ~n,
  type = "bar",
  marker = list(color = "#c45c2c"),
  hovertemplate = "Season: %{x}<br>Incidents: %{y}<extra></extra>"
) %>%
  layout(title = "Season", xaxis = list(title = ""), yaxis = list(title = "Incident count"))

p_month <- plot_ly(
  month_counts,
  x = ~Month,
  y = ~n,
  type = "bar",
  marker = list(color = "#1d6f78"),
  hovertemplate = "Month: %{x}<br>Incidents: %{y}<extra></extra>"
) %>%
  layout(title = "Month", xaxis = list(title = ""), yaxis = list(title = "Incident count"))

p_day <- plot_ly(
  day_counts,
  x = ~Day,
  y = ~n,
  type = "bar",
  marker = list(color = "#7d2f16"),
  hovertemplate = "Day: %{x}<br>Incidents: %{y}<extra></extra>"
) %>%
  layout(title = "Day of Week", xaxis = list(title = ""), yaxis = list(title = "Incident count"))

p_hour <- plot_ly(
  hour_counts,
  x = ~Hour_Label,
  y = ~n,
  type = "scatter",
  mode = "lines+markers",
  line = list(color = "#8e6c52", width = 3),
  marker = list(color = "#f0c7a5", size = 8, line = list(color = "#7d2f16", width = 1)),
  hovertemplate = "Hour: %{x}<br>Incidents: %{y}<extra></extra>"
) %>%
  layout(title = "Hour of Day", xaxis = list(title = "", tickangle = -45), yaxis = list(title = "Incident count"))

subplot(p_season, p_month, p_day, p_hour, nrows = 2, margin = 0.06, titleX = TRUE, titleY = TRUE) %>%
  layout(showlegend = FALSE)
```

The distributions show that incidents are relatively balanced across seasons, but monthly and hourly patterns carry more visible structure. Early-year months and late-evening to midnight hours stand out more clearly than any single season effect, which is why those variables matter more in later visual and modeling work.
