---
title: "Interactive Visualizations"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("scripts/prepare_project_data.R")
source("scripts/build_visual_widgets.R")
project_data <- load_project_data(year = 2020)
build_visual_widgets()

widget_iframe <- function(file_name, height = 650) {
  as.character(
    htmltools::tags$iframe(
      src = file.path("widgets", file_name),
      style = sprintf("width: 100%%; height: %dpx; border: none;", height),
      loading = "lazy"
    )
  )
}
```

<div class="hero-panel">
<div class="hero-grid">
<div class="hero-copy">
<div class="hero-kicker">Interactive EDA</div>

# Interactive Visualizations

<p class="hero-lead">These views are designed to make the main crime patterns easier to compare across neighbourhoods, premise types, and time windows. The emphasis is on robbery share, concentration, and context so the charts answer clearer public-facing questions about where and when robbery risk is highest.</p>
</div>

<div class="hero-meta">
<div class="stat-card">
<span class="stat-value">77.0%</span>
<p class="stat-label">robbery share in transit incidents, the highest among premise types</p>
</div>
<div class="stat-card">
<span class="stat-value">22.7%</span>
<p class="stat-label">peak monthly robbery share in February, versus 12.8% in June</p>
</div>
<div class="stat-card">
<span class="stat-value">15:00</span>
<p class="stat-label">hour with the highest robbery share across the Toronto incidents analyzed here</p>
</div>
</div>
</div>
</div>

The plots on this page are pre-rendered into standalone widget files in `docs/widgets`, and this page lazily embeds them for faster first-load performance.

<div class="note-callout">
The simple <code>Darkness</code> split is not a strong overall separator in this dataset. Premise type, neighbourhood concentration, month, and hour explain more of the visible variation, so the charts focus on those dimensions.
</div>

# Figure 1: Neighbourhood Concentration of Robbery Risk

**Plotly**

This chart shows median income on the x-axis, robbery share on the y-axis, rental concentration in bubble size, and robbery count in colour. Together, those dimensions show where robbery pressure is most concentrated across Toronto neighbourhoods.

```{r fig1, results='asis'}
cat(widget_iframe("figure1.html", 700))
```

# Figure 2: Premise Type vs. Daylight and Dark-Hour Robbery Share

**Plotly**

This view compares robbery share directly within each premise type. It shows that premise type is the dominant separator and that the day versus dark-hour gap is modest overall, with only a few premise types shifting materially.

```{r fig2, results='asis'}
cat(widget_iframe("figure2.html", 700))
```

# Figure 3: Premise-Type Risk Across Time {.tabset}

**Plotly**

These charts now use robbery share instead of raw robbery count, which makes premise-to-premise comparisons more defensible. The hourly panel is shown as a heatmap because multiple line traces over 24 hours were too dense to parse quickly.

## Season
```{r fig3-season, results='asis'}
cat(widget_iframe("figure3_season.html", 650))
```

## Month
```{r fig3-month, results='asis'}
cat(widget_iframe("figure3_month.html", 650))
```

## Day of Week
```{r fig3-day, results='asis'}
cat(widget_iframe("figure3_day.html", 650))
```

## Hour of Day
```{r fig3-hour, results='asis'}
cat(widget_iframe("figure3_hour.html", 650))
```

# Figure 4: Robberies Inside the Broader Incident Mix {.tabset}

**Plotly**

These overlays preserve the broader incident context by comparing all retained serious incidents against the robbery subset. They help distinguish periods with heavy activity from periods where robberies form an unusually large share of activity.

## Season
```{r fig4-season, results='asis'}
cat(widget_iframe("figure4_season.html", 650))
```

## Month
```{r fig4-month, results='asis'}
cat(widget_iframe("figure4_month.html", 650))
```

## Day of Week
```{r fig4-day, results='asis'}
cat(widget_iframe("figure4_day.html", 650))
```

## Hour of Day
```{r fig4-hour, results='asis'}
cat(widget_iframe("figure4_hour.html", 650))
```

# Figure 5: Neighbourhood Robbery Burden Map

**Leaflet**

This map aggregates incidents to neighbourhood centroids, uses marker size for robbery count, and uses colour for robbery share, so concentration and relative burden are visible at the same time.

```{r fig5, results='asis'}
cat(widget_iframe("figure5_map.html", 760))
```
