---
title: "Interactive Visualizations"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("scripts/prepare_project_data.R")
source("scripts/build_visual_widgets.R")
project_data <- load_project_data(year = 2020)
build_visual_widgets()

widget_iframe <- function(file_name, height = 650) {
  as.character(
    htmltools::tags$iframe(
      src = file.path("widgets", file_name),
      style = sprintf("width: 100%%; height: %dpx; border: none;", height),
      loading = "lazy"
    )
  )
}
```

<div class="hero-panel">
<div class="hero-grid">
<div class="hero-copy">
<div class="hero-kicker">Interactive EDA</div>

# Interactive Visualizations

<p class="hero-lead">These views were rebuilt around comparison quality instead of novelty. The original page leaned on 3D perspective, repeated neighbourhood counts, and raw-volume comparisons that made the visual story weaker than the underlying data. The revised widgets emphasize robbery share, concentration, and context.</p>
</div>

<div class="hero-meta">
<div class="stat-card">
<span class="stat-value">77.0%</span>
<p class="stat-label">robbery share in transit incidents, the highest among premise types</p>
</div>
<div class="stat-card">
<span class="stat-value">22.7%</span>
<p class="stat-label">peak monthly robbery share in February, versus 12.8% in June</p>
</div>
<div class="stat-card">
<span class="stat-value">15:00</span>
<p class="stat-label">hour with the highest robbery share across the retained incident set</p>
</div>
</div>
</div>
</div>

The plots on this page are pre-rendered into standalone widget files in `docs/widgets`, and this page lazily embeds them for faster first-load performance.

<div class="note-callout">
Diagnostic summary: the simple <code>Darkness</code> split is not a strong overall discriminator in this data. Premise type, neighbourhood concentration, month, and hour explain more of the visible variation, so the revised visuals prioritize those dimensions.
</div>

# Figure 1: Neighbourhood Concentration of Robbery Risk

**Plotly**

The original 3D scatter plot made it difficult to compare neighbourhoods because perspective distortion and camera angle competed with the actual signal. This replacement shows median income on the x-axis, robbery share on the y-axis, rental concentration in bubble size, and robbery count in colour. The result is a clearer view of where robbery pressure is concentrated.

```{r fig1, results='asis'}
cat(widget_iframe("figure1.html", 700))
```

# Figure 2: Premise Type vs. Daylight and Dark-Hour Robbery Share

**Plotly**

The original boxplots were analytically misleading because they repeated the same neighbourhood totals across many rows. The corrected view compares robbery share directly within each premise type. It shows that premise type is the dominant separator and that the day versus dark-hour gap is modest overall, with only a few premise types shifting materially.

```{r fig2, results='asis'}
cat(widget_iframe("figure2.html", 700))
```

# Figure 3: Premise-Type Risk Across Time {.tabset}

**Plotly**

These charts now use robbery share instead of raw robbery count, which makes premise-to-premise comparisons more defensible. The hourly panel is shown as a heatmap because multiple line traces over 24 hours were too dense to parse quickly.

## Season
```{r fig3-season, results='asis'}
cat(widget_iframe("figure3_season.html", 650))
```

## Month
```{r fig3-month, results='asis'}
cat(widget_iframe("figure3_month.html", 650))
```

## Day of Week
```{r fig3-day, results='asis'}
cat(widget_iframe("figure3_day.html", 650))
```

## Hour of Day
```{r fig3-hour, results='asis'}
cat(widget_iframe("figure3_hour.html", 650))
```

# Figure 4: Robberies Inside the Broader Incident Mix {.tabset}

**Plotly**

These overlays preserve the broader incident context by comparing all retained serious incidents against the robbery subset. They help distinguish periods with heavy activity from periods where robberies form an unusually large share of activity.

## Season
```{r fig4-season, results='asis'}
cat(widget_iframe("figure4_season.html", 650))
```

## Month
```{r fig4-month, results='asis'}
cat(widget_iframe("figure4_month.html", 650))
```

## Day of Week
```{r fig4-day, results='asis'}
cat(widget_iframe("figure4_day.html", 650))
```

## Hour of Day
```{r fig4-hour, results='asis'}
cat(widget_iframe("figure4_hour.html", 650))
```

# Figure 5: Neighbourhood Robbery Burden Map

**Leaflet**

Plotting every robbery incident as a point created downtown clutter and obscured neighbourhood-level interpretation. The revised map aggregates incidents to neighbourhood centroids, uses marker size for robbery count, and uses colour for robbery share, so concentration and relative burden are visible at the same time.

```{r fig5, results='asis'}
cat(widget_iframe("figure5_map.html", 760))
```
